Type-Safe Modular Parsing

### 2017-06-08 SLE_17/sections/RelatedWork.tex
Revision
Related work.

\paragraph{Syntactically Extensible Parsing}
Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}
are a mainstream area of modular syntax and parsing. They allow users to write
modular grammars, where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
and the parsing code. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking.

@@

\paragraph{Extensible Parsing Algorithm}
\textit{Parse table composition}~\cite{bravenboer2008parse}
is an approach where grammars are compiled to
modular parse tables. Those parse tables are expressed as DFAs
or NFAs, and later they can be composed by an algorithm, to provide
separate compilation for parsing. The generation of parse tables can
be quite expensive in terms of performance. The approach
is quite different from ours, since it uses parse
 tables, whereas we use parser combinators.
Our approach supports both
\emph{separate compilation} as well as \emph{modular
  type-checking}. Moreover, the extensibility of parsing is further
available at language composition and lexical level.

>>>>

\paragraph{Syntactically Extensible Parsing}
Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016}
are a mainstream area of modular syntax and parsing. They allow users to write
modular grammars, where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
and the parsing code. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking. It is worth mentioning that in~\cite{Viera2012}, users can define
grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
by a typed parser generator. Nevertheless, it only support separate compilation
on grammar fragments; a whole compilation is still needed for an extended grammar. And it requires
a lot of advanced language features, making client code less expressive. Our approach is simple from a
straightforward use of OO programming, and makes parsing code directly reusable.

@@

\paragraph{Extensible Parsing Algorithms}
\textit{Parse table composition}~\cite{bravenboer2008parse}
is an approach where grammars are compiled to
modular parse tables. Those parse tables are expressed as DFAs
or NFAs, and later they can be composed by an algorithm, to provide
separate compilation for parsing. The generation of parse tables can
be quite expensive in terms of performance. The approach
is quite different from ours, since it uses parse
 tables, whereas we use parser combinators.
Our approach supports both
\emph{separate compilation} as well as \emph{modular
  type-checking}. Moreover, the extensibility of parsing is further
available at language composition and lexical level.
